\documentclass[11pt]{article}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage{aliascnt}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage[nameinlink,noabbrev]{cleveref}
\usepackage{geometry}
\usepackage{array,longtable,booktabs,multirow}
\usepackage{tikz-cd}
\geometry{margin=1in}

\newcommand{\repourl}{\url{https://github.com/HiroSakuraba/DelPezzo7}}

\title{A Computer-Assisted Bound of Seven Singular Points for Rank-One klt del Pezzo Surfaces in Characteristic Three}
\author{Benjamin John Schulz\\[4pt]\small\textit{Independent researcher}}
\date{}

\newtheorem{theorem}{Theorem}[section]
\newaliascnt{proposition}{theorem}
\newtheorem{proposition}[proposition]{Proposition}
\aliascntresetthe{proposition}
\newaliascnt{lemma}{theorem}
\newtheorem{lemma}[lemma]{Lemma}
\aliascntresetthe{lemma}
\newaliascnt{corollary}{theorem}
\newtheorem{corollary}[corollary]{Corollary}
\aliascntresetthe{corollary}
\newaliascnt{claim}{theorem}
\newtheorem{claim}[claim]{Claim}
\aliascntresetthe{claim}
\theoremstyle{definition}
\newaliascnt{remark}{theorem}
\newtheorem{remark}[remark]{Remark}
\aliascntresetthe{remark}
\newaliascnt{notation}{theorem}
\newtheorem{notation}[notation]{Notation}
\aliascntresetthe{notation}
\newaliascnt{definition}{theorem}
\newtheorem{definition}[definition]{Definition}
\aliascntresetthe{definition}

\crefname{theorem}{theorem}{theorems}
\Crefname{theorem}{Theorem}{Theorems}
\crefname{proposition}{proposition}{propositions}
\Crefname{proposition}{Proposition}{Propositions}
\crefname{lemma}{lemma}{lemmas}
\Crefname{lemma}{Lemma}{Lemmas}
\crefname{corollary}{corollary}{corollaries}
\Crefname{corollary}{Corollary}{Corollaries}
\crefname{remark}{remark}{remarks}
\Crefname{remark}{Remark}{Remarks}
\crefname{definition}{definition}{definitions}
\Crefname{definition}{Definition}{Definitions}
\crefname{notation}{notation}{notations}
\Crefname{notation}{Notation}{Notations}

\newcommand{\A}{\mathbb A}
\newcommand{\Pbb}{\mathbb P}
\newcommand{\Fbb}{\mathbb F}
\newcommand{\Qbb}{\mathbb Q}
\newcommand{\Zbb}{\mathbb Z}
\newcommand{\NS}{\mathrm{NS}}
\newcommand{\rk}{\mathrm{rk}}
\newcommand{\Bl}{\mathrm{Bl}}
\newcommand{\Exc}{\mathrm{Exc}}
\newcommand{\Spec}{\mathrm{Spec}}
\newcommand{\Pic}{\mathrm{Pic}}
\newcommand{\Cl}{\mathrm{Cl}}
\newcommand{\Supp}{\mathrm{Supp}}
\newcommand{\Proj}{\mathrm{Proj}}
\newcommand{\Aut}{\mathrm{Aut}}
\newcommand{\ord}{\mathrm{ord}}
\newcommand{\mult}{\mathrm{mult}}
\newcommand{\eff}{\mathrm{eff}}
\newcommand{\disc}{\mathrm{disc}}
\newcommand{\ExcRank}{\mathrm{exc\text{-}rk}}
\newcommand{\ie}{\emph{i.e.}}
\newcommand{\eg}{\emph{e.g.}}

\begin{document}
\maketitle

\begin{abstract}
We prove that a rank-one klt del Pezzo surface over an algebraically closed field of characteristic $3$ has at most seven singular points. The formal proof is basket-first. Assuming $N(X)=8$, we carry out a finite exact basket reduction: after discriminant, odd-lattice, and anticanonical-complement screens, only five candidate baskets remain. Three baskets of type $7A_1+$ one exotic cyclic quotient singularity are then excluded by Weyl normalization of the exotic block together with a sharp bound on the orthogonal root subsystem. The mixed basket
\[
\mathcal B_{\mathrm{mix}}=\left\{\tfrac12(1,1),\ 5\times\tfrac13(1,1),\ 2\times\tfrac13(1,2)\right\}
\]
is reduced to two polarization orbits. One orbit is eliminated by an exact root-combinatorial obstruction. The surviving orbit is globally normalized to a single canonical model, reduced to seven grouped symmetry types, and each type is ruled out by a tangent-contact plus B\'ezout argument. Finally, the Gorenstein basket $8A_1$ is shown to force degree $1$; the anticanonical model is a finite separable double cover of $Q\cong \Pbb(1,1,2)$, and after resolving the vertex to $\mathbb F_2$ the branch divisor has the form $S+C$ with $C\in |3(S+2F)|$ disjoint from the negative section. A direct node bound shows that such a branch divisor supports at most six nodes, excluding $8A_1$. A brief contextual section records why explicit weighted, quotient, and birational search corridors all pointed toward this finite basket closure, but those exploratory lanes are not theorem dependencies. Therefore $N(X)\le 7$. The accompanying repository records the exact finite computations used in the proof.\footnote{GitHub repository bundle: \repourl.}
\end{abstract}

\tableofcontents

\section{Introduction}
Let $k$ be an algebraically closed field of characteristic $3$, and let $X$ be a rank-one klt del Pezzo surface over $k$. Write $N(X)$ for the number of singular points of $X$. The goal of this paper is to prove the following theorem.

\begin{theorem}[Main theorem]\label{thm:main}
Let $X$ be a rank-one klt del Pezzo surface over an algebraically closed field of characteristic $3$. Then
\[
N(X)\le 7.
\]
\end{theorem}

The argument developed here is a synthesis of two complementary viewpoints that emerged during the project.
On one side, an extended exploratory program examined explicit construction corridors---global quotients, weighted hypersurfaces, codimension-two weighted complete intersections, unprojection ladders, and higher-codimension Hilbert-series candidates---together with a birational ``Route C'' analysis of hypothetical high-singularity counterexamples on the minimal resolution. That geometric program was useful for search, for negative evidence, and for understanding where an eight-point example would have to hide. On the other side, the shortest formal proof turned out to be basket-first: reduce all hypothetical $8$-point surfaces to a finite exact list of baskets, then eliminate those survivors one by one.

This paper adopts the basket-first argument as its theorem spine, but it also retains a short contextual account of the exploratory geometric corridors. The point of that contextual section is explanatory rather than foundational. It records why the final proof was organized around exact basket reduction and why the remaining cases are precisely the ones attacked in the body of the paper. The actual theorem dependency chain begins with the basket reduction of \cref{sec:baskets}.

\medskip
\noindent
\textbf{Proof outline.}
The proof is by contradiction: we assume $N(X)=8$ and derive an impossibility in every case.
The argument has four layers.
\begin{enumerate}[label=\textup{(\arabic*)}]
\item\emph{Geometric context.} We briefly record the outcome of the exploratory weighted, quotient, and birational corridors that motivated the basket-first proof. In particular, the obvious explicit construction formats either fail to reach eight singular points at all, or produce surfaces with the wrong Picard number or with $-K$ merely nef. A complementary birational Route C program indicates that any benchmark counterexample should lie in a residual high-height sector. None of this is used as a theorem dependency in the formal proof, but it explains why the finite basket closure is the natural endgame.
\item\emph{Lattice-theoretic reduction.} The orbifold Noether formula, combined with discriminant and odd-lattice bilinear screens, reduces all possible $8$-point baskets to five survivors. This is the only stage at which a broad computer-assisted enumeration is used.
\item\emph{Root combinatorics.} The three exotic baskets ($7A_1$ plus one cyclic quotient) and the mixed basket are eliminated by showing that the lattice of exceptional curves on the minimal resolution is too constrained to support enough disjoint $(-2)$-curves or irreducible noncanonical exceptional curves. The key tools are Weyl/Cremona normalization, exact orthogonal root-system computations, and a B\'ezout/tangent-contact argument for irreducibility.
\item\emph{Classical algebraic geometry.} The pure Gorenstein basket $8A_1$ is eliminated by the degree-one weak del Pezzo double-cover model: the anticanonical model is a finite separable double cover of $Q\cong\Pbb(1,1,2)$, and after resolving the vertex to $\mathbb F_2$ the pulled-back branch divisor can support at most six nodes.
\end{enumerate}

The proof is therefore hybrid in origin but basket-first in formal structure. The first exact step is a finite reduction of all possible $8$-point baskets to five survivors; every later step is either a short exact lattice computation or a geometric contradiction. Because the exploratory phase went through several superseded branches, we write the paper only around the final corrected chain; older weighted, Route C, quasi-elliptic, and additive-action lanes are retained in the repository for provenance and future work, but not as theorem dependencies unless explicitly cited.

The five surviving $N=8$ baskets are
\[
8A_1,
\qquad
7A_1+\tfrac17(1,1),
\qquad
7A_1+\tfrac17(1,2),
\qquad
7A_1+\tfrac1{11}(1,2),
\]
and the mixed basket
\[
\mathcal B_{\mathrm{mix}}=\left\{\tfrac12(1,1),\ 5\times\tfrac13(1,1),\ 2\times\tfrac13(1,2)\right\}.
\]
The paper is organized around closing these five possibilities:
\begin{itemize}
\item In \cref{sec:context}, we explain how the exploratory weighted, quotient, and Route C corridors help motivate the final proof architecture.
\item In \cref{sec:exotic}, the three exotic baskets are excluded by normalizing the noncanonical block and bounding the size of the orthogonal root subsystem.
\item In \cref{sec:mixed}, the mixed basket is reduced to a single polarization orbit, globally normalized, and then eliminated by a tangent-contact contradiction.
\item In \cref{sec:pure8a1}, the pure Gorenstein basket $8A_1$ is reduced to the degree-$1$ weak del Pezzo model and excluded by a branch-node bound.
\end{itemize}

The accompanying repository is part of the proof architecture. It contains the exact scripts used to enumerate candidate baskets, compute discriminant and root data, and verify the finite orbit counts that appear in the paper. The computational steps are exact and finite; no floating-point numerics or random sampling enter the proof.

At the same time, the repository is intentionally broader than the logical core of the paper. It also preserves several exploratory branches---including ambient/deformation lanes, weighted and codimension-two scans, Route C residual notes, quasi-elliptic probes, wild $\alpha_3$/Artin--Schreier searches, and the earlier matroid/circuit screening tools---because those files are useful for provenance and for future work, whether human or machine-assisted. None of that exploratory material is used as a theorem dependency unless it is explicitly cited below.

\section{Geometric context and why the proof is basket-first}\label{sec:context}
The final proof in this paper is deliberately shorter than the full exploratory geometry that led to it. This section records the main geometric lessons of that exploratory phase and explains why they are kept as context rather than as theorem dependencies.

\subsection*{Explicit construction corridors}
Several natural construction formats were tested before the basket-first proof stabilized.
The broad conclusion was negative but informative.

\begin{enumerate}[label=\textup{(\roman*)}]
\item Tame global quotients of smooth complete intersections do not naturally reach the benchmark. In the most obvious diagonal-involution setup on a smooth $(2,2)$ complete intersection in $\Pbb^4$, the fixed-locus count already forces at most six quotient singularities.
\item The clean weighted hypersurface families that do reach eight singularities do so with the wrong Picard number. In particular, the explicit family
\[
\{f_n(x_0,x_1)+x_2x_3=0\}\subset \Pbb(a,a,p,q),\qquad p+q=an,
\]
produces exact eight-point examples only when the Picard number is visibly larger than one.
\item The most natural codimension-two weighted complete-intersection families either collapse back to hypersurface formats, fail quasismoothness, or force incompatible Fano-index arithmetic.
\item Quotient tricks and branch-unprojection ladders out of the exact-eight weighted sources do not repair the Picard-number defect. Instead they either lower the singularity count, introduce new divisibility obstructions, or leave the rank-one setting.
\item Higher-codimension Hilbert-series and weighted-cube searches do produce exotic exact-eight baskets, but again with the wrong numerical Picard number.
\end{enumerate}

The point of these negative results is not that they prove the theorem by themselves. They do not. Their role is diagnostic: they explain why the benchmark does not seem to live inside any obvious explicit weighted or quotient corridor and why a finite intrinsic obstruction becomes the right target.

\subsection*{Residual birational intuition}
A complementary line of attack studies hypothetical benchmark counterexamples birationally on the minimal resolution. In the exploratory notes this appears as the ``Route C'' program. Its moral conclusion is that, once the low-height and elliptic-descendant corridors are excluded, any hypothetical $8$-point example should lie in a residual high-height sector with very constrained bridge, packet, and multisection geometry.

That birational program is valuable for understanding the landscape: it explains why the remaining counterexample geometry would have to be unusually rigid, and it supplied several useful heuristic sieves during the search. But as a formal theorem strategy it is much longer than necessary for the present result. Once the exact basket reduction was in hand, the theorem could be proved more directly by eliminating five survivors through explicit lattice and surface arguments.

\subsection*{Why we do not use the corridor analysis as a theorem dependency}
There are two reasons.
First, the basket reduction already compresses the contradiction problem to a finite exact list, after which the remaining geometry is concrete and self-contained. Second, the exploratory corridor analysis is broader than the present theorem: it mixes search, diagnosis, and longer birational reductions that are useful for understanding the problem but not logically needed once the five-basket list is known.

For that reason the body of the paper is organized as follows. We keep the corridor analysis only as explanatory context, and we make the exact basket reduction of \cref{sec:baskets} the first load-bearing step of the proof.

\section{Preliminaries}\label{sec:prelim}

\begin{notation}
Throughout, $k$ is an algebraically closed field of characteristic $3$. Let
\[
f\colon Y\to X
\]
be the minimal resolution of a rank-one klt del Pezzo surface $X$. For a cyclic quotient singularity we use the notation $\frac1r(1,q)$. Whenever we use a blowup model, we write
\[
Y\cong \Bl_{p_1,\dots,p_n}\Pbb^2
\]
with basis $H,E_1,\dots,E_n$ and canonical class
\[
K_Y=-3H+E_1+\cdots+E_n.
\]
If $D=dH-\sum m_iE_i$, then $D^2=d^2-\sum m_i^2$ and $K_Y\cdot D=-3d+\sum m_i$.
\end{notation}

\begin{proposition}[Basic rational-surface facts]\label{prop:basic}
With notation as above:
\begin{enumerate}[label=\textup{(\roman*)}]
\item $Y$ is a smooth rational surface.
\item If $R\subset \NS(Y)$ is the lattice generated by the exceptional curves of $f$, then
\[
\rho(Y)=\rho(X)+\rk(R).
\]
\item If all singularities of $X$ are Du Val, then $f$ is crepant and $K_Y=f^*K_X$.
\item For every smooth rational surface $Y$,
\[
\rho(Y)=10-K_Y^2.
\]
\end{enumerate}
\end{proposition}

\begin{proof}
The rationality of $Y$ is standard for the minimal resolution of a rank-one del Pezzo surface. The Picard-number formula is immediate from the decomposition of $\NS(Y)$ into the pullback of $\NS(X)$ and the exceptional lattice. Crepancy for Du Val singularities is classical surface theory. Finally, for a smooth rational surface,
\[
\chi(\mathcal O_Y)=1,
\qquad
K_Y^2+\rho(Y)=10,
\]
by Noether's formula and the vanishing of $q(Y)$ and $p_g(Y)$.
\end{proof}

\begin{remark}
For later use we note that if $X$ is Gorenstein with only Du Val singularities, then the graded anticanonical ring is preserved by the minimal resolution. The precise statement appears in \cref{prop:ring-equality}.
\end{remark}

\begin{lemma}[Cohomological vanishing for anti-pluricanonical divisors on degree-one weak del Pezzo surfaces]\label{lem:vanishing}
Let $Y$ be a smooth weak del Pezzo surface of degree one over an algebraically closed field of characteristic different from $2$. Then for every $m\ge 1$,
\[
H^1(Y,\mathcal O_Y(-mK_Y))=0.
\]
In particular, $h^0(-mK_Y)=\chi(-mK_Y)=\frac{m(m+1)}{2}+1$.
\end{lemma}

\begin{proof}
The proof is by induction on $m$.

\medskip
\noindent
\textit{Base case $m=1$.}
By Riemann--Roch on the smooth rational surface $Y$,
\[
\chi(-K_Y)=\frac{(-K_Y)^2+(-K_Y)\cdot(-K_Y)}{2}+\chi(\mathcal O_Y)=\frac{1+1}{2}+1=2.
\]
By Serre duality, $H^2(Y,\mathcal O_Y(-K_Y))\cong H^0(Y,\mathcal O_Y(2K_Y))^{\vee}$. Since $-K_Y$ is nef and $(2K_Y)\cdot(-K_Y)=-2<0$, the divisor $2K_Y$ is not effective, so $h^2(-K_Y)=0$. Hence $h^0(-K_Y)-h^1(-K_Y)=2$ and $h^0(-K_Y)\ge 2$.

For the reverse bound, consider the exact sequence
\[
0\to \mathcal O_Y\to \mathcal O_Y(-K_Y)\to \mathcal O_D(-K_Y|_D)\to 0
\]
for any effective divisor $D\in |-K_Y|$. The curve $D$ has arithmetic genus $p_a(D)=1$ by adjunction: $(K_Y+D)\cdot D=0$. The restriction $-K_Y|_D$ has degree $(-K_Y)\cdot D=1$, so $h^0(D,\mathcal O_D(-K_Y|_D))\le 1$ by Riemann--Roch on the Gorenstein curve $D$ (whether $D$ is smooth, nodal, cuspidal, or reducible). Since $H^1(Y,\mathcal O_Y)=0$ for the rational surface $Y$, the long exact sequence gives
\[
0=H^1(Y,\mathcal O_Y)\to H^1(Y,\mathcal O_Y(-K_Y))\to H^1(D,\mathcal O_D(-K_Y|_D)).
\]
It therefore suffices to show $H^1(D,\mathcal O_D(-K_Y|_D))=0$. By Serre duality on the Gorenstein curve $D$ with $\omega_D\cong \mathcal O_D$ (from adjunction), one has $H^1(D,\mathcal O_D(-K_Y|_D))\cong H^0(D,\mathcal O_D(K_Y|_D))^{\vee}$. But $K_Y|_D$ has degree $K_Y\cdot D=-1<0$, so $h^0(D,\mathcal O_D(K_Y|_D))=0$. Therefore $h^1(-K_Y)=0$.

\medskip
\noindent
\textit{Inductive step.}
Fix any effective divisor $D\in |-K_Y|$. The curve $D$ is a Gorenstein curve of arithmetic genus $1$ with $\omega_D\cong\mathcal O_D$, but $D$ need not be smooth (in characteristic $3$ the generic member could be cuspidal if the pencil is quasi-elliptic). This does not affect the argument: we only use Serre duality on the Gorenstein curve $D$. Consider the exact sequence
\[
0\to \mathcal O_Y(-(m-1)K_Y)\to \mathcal O_Y(-mK_Y)\to \mathcal O_D(-mK_Y|_D)\to 0.
\]
The restriction $-mK_Y|_D$ has degree $(-mK_Y)\cdot(-K_Y)=m$. By Serre duality on $D$,
\[
H^1(D,\mathcal O_D(-mK_Y|_D))\cong H^0(D,\mathcal O_D(mK_Y|_D))^{\vee}.
\]
Since $mK_Y|_D$ has degree $-m<0$, we get $h^0(D,\mathcal O_D(mK_Y|_D))=0$, hence $H^1(D,\mathcal O_D(-mK_Y|_D))=0$. The long exact sequence in cohomology gives
\[
H^1(Y,\mathcal O_Y(-(m-1)K_Y))\to H^1(Y,\mathcal O_Y(-mK_Y))\to H^1(D,\mathcal O_D(-mK_Y|_D))=0.
\]
By the inductive hypothesis, $H^1(Y,\mathcal O_Y(-(m-1)K_Y))=0$. Therefore $H^1(Y,\mathcal O_Y(-mK_Y))=0$.

The formula $h^0=\chi$ then follows from $H^2(-mK_Y)\cong H^0((m+1)K_Y)^{\vee}=0$.
\end{proof}

\section{Reduction to five candidate baskets}\label{sec:baskets}

This section is the only place where the proof uses a broad finite computer-assisted search. The underlying computation is exact. It proceeds in four stages:
\begin{enumerate}[label=\textup{(\arabic*)}]
\item enumerate all local cyclic quotient types that survive the elementary $N=8$ local constraints,
\item enumerate all $8$-point multisets built from those local types,
\item impose orbifold and discriminant constraints,
\item replace the provisional quadratic-form screen by the corrected odd-lattice screen on the discriminant bilinear form and anticanonical complement.
\end{enumerate}
The output is the five-basket list of \cref{prop:final-baskets}.

\begin{proposition}[Final $N=8$ basket reduction]\label{prop:final-baskets}
Assume $X$ is a rank-one klt del Pezzo surface over $k$ with $N(X)=8$. Then the singularity basket of $X$ is one of the following five baskets:
\[
8A_1,
\qquad
7A_1+\tfrac17(1,1),
\qquad
7A_1+\tfrac17(1,2),
\qquad
7A_1+\tfrac1{11}(1,2),
\]
and
\[
\mathcal B_{\mathrm{mix}}=\left\{\tfrac12(1,1),\ 5\times\tfrac13(1,1),\ 2\times\tfrac13(1,2)\right\}.
\]
\end{proposition}

\begin{proof}
We summarize the exact finite reduction; the full local-type table, scripts, and machine-readable outputs appear in \cref{app:basket,app:repo}. The reduction is exact and finite at every stage.

The local screen leaves $17$ cyclic quotient types, namely
\[
\tfrac12(1,1),\ \tfrac13(1,1),\ \tfrac14(1,1),\ \tfrac15(1,1),\ \tfrac16(1,1),\ \tfrac17(1,1),\ \tfrac18(1,1),\ \tfrac19(1,1),\ \tfrac1{10}(1,1),\ \tfrac1{11}(1,1),
\]
\[
\tfrac13(1,2),\ \tfrac15(1,2),\ \tfrac18(1,3),\ \tfrac17(1,2),\ \tfrac1{11}(1,3),\ \tfrac19(1,2),\ \tfrac1{11}(1,2).
\]
Exact enumeration of all $8$-point multisets built from these types gives $735{,}471$ baskets. The successive exact reductions are summarized in \cref{tab:basketcounts}.

\begin{table}[h]
\centering
\begin{tabular}{l r}
\toprule
stage & number of baskets \\
\midrule
all $8$-point multisets from the $17$ local types & $735{,}471$ \\
positive candidate $K_X^2$ & $732{,}878$ \\
coarse square-discriminant budget & $4{,}894$ \\
corrected exact small-discriminant window (resolved) & $45$ \\
final survivors after odd-lattice bilinear / anticanonical screen & $5$ \\
\bottomrule
\end{tabular}
\caption{Exact basket-reduction counts used in \cref{prop:final-baskets}.}
\label{tab:basketcounts}
\end{table}

In the small-discriminant window, the corrected exact search resolves $45$ baskets in total. Of these, $24$ survive the preliminary exact discriminant stage and $21$ are eliminated there. In the earlier project history, two of those $45$ had temporarily remained unresolved in a provisional pipeline; the corrected pipeline used here resolves both of them negatively. After the final odd-lattice bilinear/anticanonical screen, exactly the five baskets listed above remain.

The proof is exact and finite. The scripts, machine-readable outputs, and intermediate counts are recorded in the repository and summarized in \cref{app:basket,app:repo}.
\end{proof}

\begin{lemma}[How the numerical basket screens are formed]\label{lem:basket-screens}
Let $\mathcal B=\{P_1,\dots,P_8\}$ be an $8$-point basket drawn from the local types of \cref{tab:localtypes}. For each point $P_i$, let $r_i$ be the rank of the exceptional chain on the minimal resolution and let $\delta_i$ denote the local contribution of $P_i$ to the global correction $K_X^2-K_Y^2$, as recorded in \cref{tab:localtypes}. Then any rank-one klt del Pezzo surface with basket $\mathcal B$ must satisfy
\[
K_X^2 = 9-\sum_{i=1}^8 r_i + \sum_{i=1}^8 \delta_i > 0.
\]
Moreover, if $R$ denotes the total exceptional lattice and $m$ is any positive integer for which $mK_X$ is Cartier, then the discriminant compatibility for the rank-one complement in $\NS(Y)$ imposes the square condition
\[
|A_R|\,m^2K_X^2\in (\mathbf Z_{>0})^2,
\]
and the refined odd-lattice screen requires that the discriminant bilinear form of $R$ admit a cyclic rank-one complement carrying the anticanonical class of the required norm.
\end{lemma}

\begin{proof}
Since $\rho(X)=1$ and the total exceptional rank is $r:=\sum r_i$, one has
\[
\rho(Y)=1+r.
\]
Because $Y$ is smooth and rational, \cref{prop:basic}(iv) gives
\[
K_Y^2 = 10-\rho(Y)=9-r.
\]
By definition, each local correction $\delta_i$ satisfies $K_X^2-K_Y^2=\delta_i$ across the corresponding singularity, so summing over the basket gives
\[
K_X^2 = K_Y^2+\sum_{i=1}^8 \delta_i = 9-\sum r_i+\sum \delta_i.
\]
Positivity of $-K_X$ forces $K_X^2>0$, giving the first screen.

For the second statement, write $\bar R\subset \NS(Y)$ for the primitive closure of the exceptional lattice. Since $\NS(Y)$ is an odd unimodular lattice of signature $(1,\rho(Y)-1)$, a rank-one complement generated by a class $h$ of square $h^2=m^2K_X^2$ can exist only if the discriminant group of $\bar R\oplus \mathbf Zh$ admits an isotropic glue to a unimodular overlattice. In particular the product of discriminants must be a square, which yields the displayed necessary condition. The corrected final screen is the exact version of this statement: one works prime-by-prime with the discriminant bilinear form of $\bar R$ and asks whether there is a cyclic rank-one complement carrying the anticanonical class of the required norm. This is the odd-lattice bilinear/anticanonical test implemented in the repository.
\end{proof}

\begin{remark}[Computer-assisted status]\label{rem:computer-assisted}
\Cref{prop:final-baskets} is the only proposition in the paper that depends on a broad exact enumeration rather than a short intrinsic geometric argument. All later eliminations are either hand proofs or exact finite checks attached to a single already-normalized configuration.
\end{remark}

\section{The three exotic baskets}\label{sec:exotic}
We first eliminate the three baskets
\[
7A_1+\tfrac17(1,1),\qquad 7A_1+\tfrac17(1,2),\qquad 7A_1+\tfrac1{11}(1,2).
\]

\subsection{Normalization of the noncanonical block}

\begin{proposition}[Standard-form normalization]\label{prop:exotic-normalization}
Let $Y\to X$ be the minimal resolution of one of the three exotic baskets above. Then, up to Weyl/Cremona action, the noncanonical exceptional block admits a unique standard representative:
\begin{align*}
\tfrac17(1,1) &:\quad C\sim H-E_1-\cdots-E_8,\\
\tfrac17(1,2) &:\quad C_2=E_5-E_6,\qquad C_1\sim H-E_1-\cdots-E_5,\\
\tfrac1{11}(1,2) &:\quad C_2=E_7-E_8,\qquad C_1\sim H-E_1-\cdots-E_7.
\end{align*}
Here the chains are ordered so that the $(-2)$-component is denoted by $C_2$ whenever the exotic point resolves to a two-curve chain.
\end{proposition}

\begin{proof}
We solve the standard-form equations in each case.

\medskip
\noindent
\textbf{Case $7A_1+\frac17(1,1)$.}
The noncanonical exceptional curve $C$ satisfies
\[
C^2=-7,\qquad K_Y\cdot C=5.
\]
The total exceptional rank is $7\cdot 1+1=8$ (seven $A_1$-curves of rank $1$ each, plus the single noncanonical curve), so $\rho(Y)=1+8=9$ and $K_Y^2=10-9=1$. Write
\[
C=dH-\sum_{i=1}^8m_iE_i
\]
in standard form with $d\ge m_1\ge \cdots\ge m_8\ge 0$. Then
\[
\sum m_i=3d+5,\qquad \sum m_i^2=d^2+7.
\]
By Cauchy,
\[
(3d+5)^2\le 8(d^2+7),
\]
which forces $d\le 1$. Since $d=0$ is impossible for an effective standard-form class of self-intersection $-7$, we get $d=1$. Then
\[
\sum m_i=8,\qquad \sum m_i^2=8,
\]
so necessarily $m_1=\cdots=m_8=1$.

\medskip
\noindent
\textbf{Case $7A_1+\frac17(1,2)$.}
The exceptional chain is $[4,2]$, which has rank $2$. The total exceptional rank is $7\cdot 1+2=9$, giving $\rho(Y)=10$ and $K_Y^2=0$, so $Y=\Bl_9\Pbb^2$. The $(-2)$-curve $C_2$ is a real root in the affine $E_8$ lattice $K_Y^{\perp}/\Zbb K_Y$, so the Weyl group is transitive on its root orbit. We may therefore normalize
\[
C_2=E_5-E_6.
\]
Write
\[
C_1=dH-\sum_{i=1}^9m_iE_i
\]
in standard form. Then
\[
\sum m_i=3d+2,\qquad \sum m_i^2=d^2+4,\qquad m_5-m_6=1.
\]
The exact standard-form solve for the first two equations has only two solutions:
\[
(1;1,1,1,1,1,0,0,0,0),\qquad (2;1,1,1,1,1,1,1,1,0).
\]
The second violates $m_5-m_6=1$, so the first is the unique admissible solution. Hence
\[
C_1\sim H-E_1-\cdots-E_5.
\]

\medskip
\noindent
\textbf{Case $7A_1+\frac1{11}(1,2)$.}
The exceptional chain is $[6,2]$, which has rank $2$. The total exceptional rank is $7\cdot 1+2=9$, giving $\rho(Y)=10$ and $K_Y^2=0$, so again $Y=\Bl_9\Pbb^2$. We normalize the root component to
\[
C_2=E_7-E_8.
\]
For the other component,
\[
\sum m_i=3d+4,\qquad \sum m_i^2=d^2+6,\qquad m_7-m_8=1.
\]
The exact reduced solve gives the unique solution
\[
(1;1,1,1,1,1,1,1,0,0),
\]
hence
\[
C_1\sim H-E_1-\cdots-E_7.
\]
This proves the proposition.
\end{proof}

\subsection{The orthogonal root subsystems}

\begin{lemma}[Orthogonal real-root sets for the exotic baskets]\label{lem:exotic-root-types}
In the three normal forms of \cref{prop:exotic-normalization}, the orthogonal real-root sets have cardinalities
\[
56,\qquad 48,\qquad 32.
\]
More precisely:
\begin{enumerate}[label=\textup{(\roman*)},leftmargin=2.5em]
    \item for $\frac17(1,1)$ the orthogonal root subsystem is of type $A_7$;
    \item for $\frac17(1,2)$ the orthogonal real-root set has $48$ elements and contains at most $4$ pairwise orthogonal roots;
    \item for $\frac1{11}(1,2)$ the orthogonal real-root set has $32$ elements and contains at most $4$ pairwise orthogonal roots.
\end{enumerate}
In particular, the maximal number of pairwise orthogonal roots is at most
\[
4,\qquad 4,\qquad 4.
\]
\end{lemma}

\begin{proof}
We record both the structural picture and the exact finite verification.

For $\frac17(1,1)$, the orthogonal roots to
\[
C=H-E_1-\cdots-E_8
\]
inside $K_Y^{\perp}\cong E_8(-1)$ contain the chain
\[
E_1-E_2,\ E_2-E_3,\ \dots,\ E_7-E_8,
\]
which gives a root subsystem of type $A_7$. The exact root enumeration in the repository returns exactly $56$ orthogonal real roots, which is the full root count of $A_7$; hence the orthogonal root subsystem is exactly $A_7$. In particular the maximal number of pairwise orthogonal roots is $4$.

For $\frac17(1,2)$, after normalizing
\[
C_2=E_5-E_6,\qquad C_1=H-E_1-\cdots-E_5,
\]
the orthogonal real-root set in $K_Y^{\perp}$ can be described explicitly as the union of
\begin{itemize}[leftmargin=2.5em]
    \item the $12$ permutation roots on $\{E_1,E_2,E_3,E_4\}$,
    \item the $6$ permutation roots on $\{E_7,E_8,E_9\}$,
    \item and $15$ roots of the form $H-E_a-E_b-E_c$ together with their negatives,
\end{itemize}
for a total of $12+6+30=48$ roots. An exact clique computation on the orthogonality graph shows that the largest pairwise orthogonal subset has size $4$.

For $\frac1{11}(1,2)$, after normalizing
\[
C_2=E_7-E_8,\qquad C_1=H-E_1-\cdots-E_7,
\]
the orthogonal real-root set consists of the $30$ permutation roots on $\{E_1,\dots,E_6\}$ together with the two roots
\[
\pm(H-E_7-E_8-E_9),
\]
for a total of $32$ roots. Again an exact clique computation on the orthogonality graph shows that the maximum number of pairwise orthogonal roots is $4$.
\end{proof}

\begin{theorem}[Exotic baskets are impossible]\label{thm:exotic}
None of the three baskets
\[
7A_1+\tfrac17(1,1),\qquad 7A_1+\tfrac17(1,2),\qquad 7A_1+\tfrac1{11}(1,2)
\]
occur on a rank-one klt del Pezzo surface in characteristic $3$.
\end{theorem}

\begin{proof}
Each basket would require seven further disjoint $A_1$-curves orthogonal to the noncanonical block. By \cref{lem:exotic-root-types}, the orthogonal real-root subsystem is too small in every case: it supports at most $4$, $4$, and $4$ pairwise orthogonal roots respectively. This contradiction rules out all three baskets.
\end{proof}

\section{The mixed basket}\label{sec:mixed}
We now eliminate the mixed basket
\[
\mathcal B_{\mathrm{mix}}=\left\{\tfrac12(1,1),\ 5\times\tfrac13(1,1),\ 2\times\tfrac13(1,2)\right\}.
\]
The proof is longer, but every step after the initial basket reduction is finite and exact.

\subsection{Polarization reduction}

\begin{proposition}[The mixed-basket polarization]\label{prop:mixed-polarization}
Assume $X$ has basket $\mathcal B_{\mathrm{mix}}$, and let $f\colon Y\to X$ be the minimal resolution. If $C_1,\dots,C_5$ are the five noncanonical $(-3)$-curves coming from the five $\frac13(1,1)$ points, then
\[
L:=3f^*(-K_X)=-3K_Y-\sum_{i=1}^5C_i
\]
is an integral nef divisor satisfying
\[
L^2=6,\qquad K_Y\cdot L=-2.
\]
After Cremona reduction, $L$ lies in one of exactly two standard-form orbits:
\[
L_1=4H-E_1-\cdots-E_{10},
\qquad
L_2=6H-2E_1-\cdots-2E_7-E_8-E_9.
\]
\end{proposition}

\begin{proof}
For a $\frac13(1,1)$ point, the discrepancy on the exceptional curve is $-\frac13$, while the singularities $A_1$ and $\frac13(1,2)$ are canonical. Thus
\[
K_Y=f^*K_X-\frac13\sum_{i=1}^5C_i.
\]
Rearranging gives the stated integral divisor $L$. Since $f^*(-K_X)$ is nef and big, so is $L$. Direct intersection computation yields
\[
L^2=9K_Y^2+6\sum K_Y\cdot C_i+\sum_{i,j}C_i\cdot C_j=6,
\qquad
K_Y\cdot L=-3K_Y^2-\sum K_Y\cdot C_i=-2.
\]
Write
\[
L=dH-\sum_{i=1}^{10}m_iE_i
\]
in standard form. Then
\[
\sum m_i=3d-2,
\qquad
\sum m_i^2=d^2-6.
\]
A finite reduced standard-form solve gives exactly two solutions up to Cremona equivalence, namely $L_1$ and $L_2$.
\end{proof}

\begin{proposition}[The polarization orbit $L_1$ is impossible]\label{prop:L1-impossible}
The mixed basket cannot occur in the polarization orbit $L_1$.
\end{proposition}

\begin{proof}
Define the root sets
\[
\mathcal R_0(L)=\{r\mid r^2=-2,\ K_Y\cdot r=0,\ L\cdot r=0\},
\qquad
\mathcal R_{-2}(L)=\{r\mid r^2=-2,\ K_Y\cdot r=0,\ L\cdot r=-2\}.
\]
For $L=L_1$, the exact search gives
\[
|\mathcal R_0(L_1)|=90,
\qquad
|\mathcal R_{-2}(L_1)|=210.
\]
A mixed-basket realization would require a $5$-clique in $\mathcal R_{-2}(L_1)$ with pairwise intersections $1$, together with an orthogonal canonical subsystem of type $A_2+A_2+A_1$ in $\mathcal R_0(L_1)$. The graph on $\mathcal R_{-2}(L_1)$ contains exactly $30{,}240$ such transformed noncanonical five-cliques, but the exact compatibility search finds that none is orthogonal to an $A_2+A_2+A_1$ subsystem. Hence $L_1$ is impossible.
\end{proof}

Thus any mixed-basket realization must occur in the single surviving orbit $L_2$.

\subsection{Global normalization in the \texorpdfstring{$L_2$}{L2} orbit}

\begin{theorem}[Global normalization theorem for the surviving $L_2$ orbit]\label{thm:L2-global-normalization}
Inside the $L_2$ root-theoretic model, canonical subsystems of type $A_2+A_2+A_1$ split into exactly two Weyl orbits, of sizes
\[
3360\qquad\text{and}\qquad 10080.
\]
Among these two orbits, only the orbit of size $3360$ is compatible with a transformed noncanonical $5$-clique. Consequently, every mixed-basket lattice configuration in the $L_2$ orbit is Weyl-equivalent to a canonical subsystem in the compatible $3360$-orbit. For the grouped geometric analysis below we fix the convenient representative
\[
S_{\mathrm{std}}=\langle E_3-E_1,\ E_1-E_2,\ E_6-E_4,\ E_5-E_6,\ E_7-E_8\rangle.
\]
\end{theorem}

\begin{proof}
Set
\[
\mathcal R_0=\mathcal R_0(L_2),\qquad \mathcal R_{-2}=\mathcal R_{-2}(L_2).
\]
The exact root search gives
\[
|\mathcal R_0|=128,
\qquad
|\mathcal R_{-2}|=177.
\]
Now enumerate all canonical subsystems of type $A_2+A_2+A_1$ inside $\mathcal R_0$. There are exactly
\[
13{,}440
\]
such rootsets. Under the Weyl group generated by reflections in the roots of $\mathcal R_0$, these split into two orbits of sizes $3360$ and $10080$.

Next enumerate all transformed noncanonical five-cliques inside $\mathcal R_{-2}$, 
\ie{} all $5$-subsets with pairwise intersection $1$. There are exactly
\[
5600
\]
such cliques. For each one, enumerate all orthogonal canonical subsystems of type $A_2+A_2+A_1$. The result is uniform: every transformed noncanonical five-clique is compatible only with the orbit of size $3360$, namely the orbit containing $S_{\mathrm{std}}$. None is compatible with the second canonical orbit.

Therefore every full mixed-basket lattice configuration can be moved by a $W(L_2)$-isometry so that its canonical subsystem is exactly $S_{\mathrm{std}}$.
\end{proof}

\subsection{Grouped reduction}

\begin{proposition}[Grouped reduction in the normalized $L_2$ model]\label{prop:grouped-reduction}
After the normalization of \cref{thm:L2-global-normalization}, every noncanonical class in the mixed-basket configuration has grouped form
\[
D=dH-a(E_1+E_2+E_3)-b(E_4+E_5+E_6)-c(E_7+E_8)-mE_9-nE_{10},
\]
with multiplicity equalities
\[
m_1=m_2=m_3,\qquad m_4=m_5=m_6,\qquad m_7=m_8.
\]
Up to the natural symmetries exchanging the two length-$3$ clusters and the two terminal points, the resulting pairwise-orthogonal transformed noncanonical five-cliques fall into exactly seven orbit types.
\end{proposition}

\begin{proof}
Orthogonality to the five roots spanning $S_{\mathrm{std}}$ forces the displayed equalities:
\[
(E_3-E_1)\cdot D=0\Rightarrow m_1=m_3,
\qquad
(E_1-E_2)\cdot D=0\Rightarrow m_1=m_2,
\]
\[
(E_6-E_4)\cdot D=0\Rightarrow m_4=m_6,
\qquad
(E_5-E_6)\cdot D=0\Rightarrow m_5=m_6,
\]
\[
(E_7-E_8)\cdot D=0\Rightarrow m_7=m_8.
\]
Thus every candidate noncanonical class is grouped. The exact clique enumeration in the normalized grouped model produces $16$ grouped five-cliques; quotienting by the obvious symmetries leaves precisely $7$ orbit types.
\end{proof}

\subsection{Tangent contact and the seven orbit types}

\begin{lemma}[Tangent-contact lemma]\label{lem:tangent-contact}
Let
\[
D=dH-a(E_1+E_2+E_3)-b(E_4+E_5+E_6)-c(E_7+E_8)-mE_9-nE_{10}
\]
be an effective grouped class in the normalized model. Let $T_A,T_B,T_C$ denote the tangent lines to the three proximate clusters of lengths $3,3,2$ respectively. Then
\[
I(T_A,D)\ge 3a,
\qquad
I(T_B,D)\ge 3b,
\qquad
I(T_C,D)\ge 2c.
\]
Consequently, if any one of the inequalities
\[
3a>d,
\qquad
3b>d,
\qquad
2c>d
\]
holds, then the corresponding tangent line is forced as a component by B\'ezout, and the strict transform of $D$ on the blowup is reducible.
\end{lemma}

\begin{proof}
The argument is local and characteristic-free; it uses only the theory of infinitely near points in the plane (see \eg{} \cite[\S3.3]{CasasAlvero00}). The length-$3$ multiplicity sequence $(a,a,a)$ says that the plane curve representing $D$ meets the tangent line $T_A$ with contact order at least $3a$ after successively blowing up the three infinitely near points in the $A$-cluster: at each stage the strict transform passes through the exceptional point with multiplicity at least $a$, and the total contact order is the sum. The same argument gives contact order at least $3b$ with $T_B$ and $2c$ with $T_C$. If one of these contact orders exceeds the plane degree $d$, then B\'ezout forces the corresponding tangent line to be an actual component of the plane curve. The strict transform is therefore reducible.\qedhere
\end{proof}

\begin{proposition}[The seven grouped $L_2$ orbit types are all impossible]\label{prop:seven-orbits}
For each of the seven grouped orbit types from \cref{prop:grouped-reduction}, one of the five classes in any representative grouped five-clique violates one of the inequalities in \cref{lem:tangent-contact}. Consequently every grouped five-clique in each orbit type contains a reducible member, so none of the seven orbit types can realize five irreducible noncanonical exceptional $(-3)$-curves.
\end{proposition}

\begin{proof}
For each orbit type we inspect a representative grouped five-clique and exhibit one of its five classes that is necessarily reducible. Write a grouped class as $(d;a,b,c;m,n)$. The exact orbit representatives are listed in the repository; the following table records, for each orbit type, a member of the representative five-clique that is forced reducible by \cref{lem:tangent-contact}. Because the orbit type is an orbit of whole grouped five-cliques under the residual symmetry group, the existence of such a distinguished reducible member is invariant across the orbit type.

\begin{center}
\begin{tabular}{c|c|c}
\toprule
Orbit & representative $(d;a,b,c;m,n)$ & forced inequality \\
\midrule
1 & $(5;2,1,2;1,2)$ & $3a=6>5$ \\
2 & $(2;1,1,0;0,1)$ & $3a=3>2$, $3b=3>2$ \\
3 & $(2;1,1,0;0,1)$ & $3a=3>2$, $3b=3>2$ \\
4 & $(7;3,2,2;2,1)$ & $3a=9>7$ \\
5 & $(5;1,2,2;2,1)$ & $3b=6>5$ \\
6 & $(4;1,2,1;1,1)$ & $3b=6>4$ \\
7 & $(4;1,2,1;1,1)$ & $3b=6>4$ \\
\bottomrule
\end{tabular}
\end{center}

In every case the relevant tangent line is forced as a component by \cref{lem:tangent-contact}, so the displayed member of the representative five-clique is reducible. Since grouped-orbit equivalence acts on whole five-cliques, every five-clique in the same orbit type contains a corresponding reducible member. Therefore no orbit type can realize the required five irreducible noncanonical exceptional curves. Note that orbits $2$ and $3$ share the same displayed forced-reducible class, as do orbits $6$ and $7$; the symmetry orbits are different five-cliques, but one may choose the same witness class in each pair.
\end{proof}

\begin{theorem}[The mixed basket is impossible]\label{thm:mixed-impossible}
The mixed basket
\[
\mathcal B_{\mathrm{mix}}=\left\{\tfrac12(1,1),\ 5\times\tfrac13(1,1),\ 2\times\tfrac13(1,2)\right\}
\]
does not occur on a rank-one klt del Pezzo surface in characteristic $3$.
\end{theorem}

\begin{proof}
By \cref{prop:mixed-polarization}, any realization of the mixed basket lies in one of the polarization orbits $L_1$ or $L_2$. The orbit $L_1$ is impossible by \cref{prop:L1-impossible}. Thus only $L_2$ remains.

By \cref{thm:L2-global-normalization}, every full $L_2$ configuration is Weyl-equivalent to the standard canonical subsystem $S_{\mathrm{std}}$. By \cref{prop:grouped-reduction}, every transformed noncanonical five-clique is then one of the seven grouped orbit types. By \cref{prop:seven-orbits}, every such grouped five-clique contains a reducible member, contradicting irreducibility of the five noncanonical exceptional curves. Therefore the mixed basket cannot occur.
\end{proof}

\section{The pure \texorpdfstring{$8A_1$}{8A1} case}\label{sec:pure8a1}
We now eliminate the remaining Gorenstein basket $8A_1$.

\subsection{Degree one and the weak del Pezzo model}

\begin{proposition}[The pure $8A_1$ case has degree one]\label{prop:8a1-degree-one}
Assume $X$ has basket $8A_1$. Then
\[
K_X^2=1.
\]
\end{proposition}

\begin{proof}
Let $f\colon Y\to X$ be the minimal resolution. Since the singularities are all Du Val of type $A_1$, the resolution is crepant, so
\[
K_Y=f^*K_X,
\qquad
K_Y^2=K_X^2=:d.
\]
Also
\[
\rho(Y)=\rho(X)+8=9.
\]
By \cref{prop:basic}(iv),
\[
10-K_Y^2=\rho(Y)=9,
\]
so $d=1$.
\end{proof}

\begin{definition}[Almost general position]
A set of eight points in $\Pbb^2$ is said to be in \emph{almost general position} if no three are collinear, no six lie on a conic, and no cubic passes through all eight with one of them as a singular point. Equivalently, the blowup has nef anticanonical class.
\end{definition}

\begin{remark}[Almost general position versus the presence of $(-2)$-curves]
The phrase ``almost general position'' in \cref{prop:deg1-weak-dp} is the standard weak-del-Pezzo condition: it ensures that $-K_Y$ is nef. It does \emph{not} mean that the eight points are generic in the everyday sense. In particular, the existence of $(-2)$-curves on a weak del Pezzo surface is compatible with almost general position; these curves record degeneracies in the blowup configuration that are still allowed by the nefness of $-K_Y$.
\end{remark}

\begin{proposition}[Degree-one weak del Pezzo structure]\label{prop:deg1-weak-dp}
Let $Y$ be a weak del Pezzo surface of degree one over an algebraically closed field of characteristic $3$. Then:
\begin{enumerate}[label=\textup{(\roman*)}]
\item $Y\cong \Bl_{p_1,\dots,p_8}\Pbb^2$ for eight points in almost general position;
\item $|-K_Y|$ is a pencil with exactly one base point $q$;
\item after blowing up $q$, one obtains a smooth rational surface
\[
\sigma\colon Z:=\Bl_qY\to Y
\]
whose anticanonical pencil induces a relatively minimal genus-one fibration $Z\to\Pbb^1$;
\item the linear system $|-2K_Y|$ is basepoint-free and induces a separable double cover of the quadric cone.
\end{enumerate}
\end{proposition}

\begin{proof}
Statement (i) is the standard degree-one weak-del-Pezzo classification: a weak del Pezzo surface of degree one is the blowup of $\Pbb^2$ at eight points in almost general position; see Demazure \cite{Demazure80}. Statement (ii) is the classical description of the anti-canonical system in degree one: $|-K_Y|$ is a genus-one pencil with a unique base point. Indeed, by Riemann--Roch $\chi(-K_Y)=2$, and the vanishing $H^1(Y,\mathcal O_Y(-K_Y))=0$ (from \cref{lem:vanishing}) gives $h^0(-K_Y)=2$. Blowing up that base point gives the smooth rational surface $Z$ together with the induced relatively minimal genus-one fibration $Z\to\Pbb^1$ of (iii), which in characteristic $3$ may be elliptic or quasi-elliptic. Statement (iv) is the bi-anticanonical model theorem in characteristic different from $2$: the linear system $|-2K_Y|$ is basepoint-free and induces a separable double cover of the quadric cone. For normal Gorenstein degree-one del Pezzo surfaces this is recorded by Hidaka--Watanabe \cite{HW81}; for degree-one del Pezzo surfaces with canonical singularities the same double-cover description appears in Proposition~2.2 and Remark~2.3 of Ascher--Bejleri \cite{AscherBejleri21}. The basepoint-freeness of $|-2K_Y|$ on a weak del Pezzo of degree one also uses $H^1(Y,\mathcal O_Y(-2K_Y))=0$, which follows from \cref{lem:vanishing}. The separability in characteristic $3$ is immediate because the covering degree is $2$.
\end{proof}

\begin{lemma}[The anti-canonical base point avoids the $(-2)$-curves]\label{lem:basepoint-avoids-minus2}
Let $Y$ be a degree-one weak del Pezzo surface, and let $q$ be the unique base point of $|-K_Y|$. Then $q$ lies on no $(-2)$-curve of $Y$.
\end{lemma}

\begin{proof}
Let $R\subset Y$ be a $(-2)$-curve. Since $-K_Y$ is nef and $R$ is orthogonal to the anti-canonical class, one has
\[
(-K_Y)\cdot R=0.
\]
Assume for contradiction that $q\in R$. Every divisor $D\in |-K_Y|$ passes through $q$, so the effective divisors $D$ and $R$ meet. Because $D\cdot R=0$, positivity of local intersection numbers forces $R$ to be a component of every $D$. Hence $R$ is a fixed component of $|-K_Y|$.

Write
\[
|-K_Y| = R + |M|.
\]
Then $M\sim -K_Y-R$, and since $h^0(-K_Y)=2$ (by Riemann--Roch and \cref{lem:vanishing}, as $-K_Y$ is nef), the moving part $|M|$ is nonempty. But
\[
M^2 = (-K_Y-R)^2 = K_Y^2 + R^2 - 2(-K_Y\cdot R)=1-2-0=-1,
\]
which is impossible for the moving part of a nontrivial pencil. Therefore $q\notin R$.
\end{proof}

\begin{proposition}[Equality of anticanonical rings under the crepant resolution]\label{prop:ring-equality}
Let $X$ be a Gorenstein del Pezzo surface with only Du Val singularities, and let $f\colon Y\to X$ be its minimal resolution. Then for every $m\ge 0$,
\[
H^0(X,\mathcal O_X(-mK_X))\cong H^0(Y,\mathcal O_Y(-mK_Y)),
\]
so
\[
R(X,-K_X)\cong R(Y,-K_Y).
\]
\end{proposition}

\begin{proof}
Du Val singularities are rational, so $R^if_*\mathcal O_Y=0$ for $i>0$ and $f_*\mathcal O_Y=\mathcal O_X$. Since $f$ is crepant, $K_Y=f^*K_X$, hence
\[
\mathcal O_Y(-mK_Y)\cong f^*\mathcal O_X(-mK_X)
\]
for every $m\ge 0$. By the projection formula,
\[
f_*\mathcal O_Y(-mK_Y)\cong \mathcal O_X(-mK_X)\otimes f_*\mathcal O_Y\cong \mathcal O_X(-mK_X).
\]
Taking global sections proves the claim.
\end{proof}

\subsection{The separable branch model}

The key structural point is that the anticanonical model of $X$ (equivalently of $Y$, by \cref{prop:ring-equality}) is a finite separable double cover of the quadric cone. To apply the node correspondence, we work with this finite cover directly, not with the smooth blowup $Z$.

\begin{proposition}[The anticanonical model as a finite double cover]\label{prop:branch-model}
Assume $X$ has basket $8A_1$, and let $f\colon Y\to X$ be the minimal resolution. Then the morphism induced by $|-2K_Y|$ factors as
\[
Y\xrightarrow{f} X \xrightarrow{\pi} Q\cong \Pbb(1,1,2),
\]
where $f$ is the minimal resolution and $\pi$ is a finite separable double cover. After resolving the vertex of $Q$ to obtain $\rho\colon \mathbb F_2\to Q$, the normalized fiber product gives the commutative diagram
\[
\begin{tikzcd}
Z \ar[r,"\sigma"] \ar[d,"\widetilde\pi"'] & Y \ar[r,"f"] & X \ar[d,"\pi"] \\
\mathbb F_2 \ar[rr,"\rho"] & & Q
\end{tikzcd}
\]
where $Z=\Bl_qY$ is smooth, $\widetilde\pi\colon Z\to \mathbb F_2$ is a generically finite morphism of degree $2$ (but not finite: it contracts the strict transforms of the eight $(-2)$-curves to points), and $\pi\colon X\to Q$ is the finite separable double cover branched along a weighted sextic.

On $\mathbb F_2$, the branch divisor of the finite cover $\pi$ pulls back to
\[
B=S+C,
\]
where $S$ is the negative section and $C\in |3M|$ is a reduced divisor, $M:=S+2F$, $C\cdot S=0$.
\end{proposition}

\begin{proof}
By \cref{prop:deg1-weak-dp}(iv), $|-2K_Y|$ is basepoint-free and induces a separable morphism of degree $2$ to $Q\cong\Pbb(1,1,2)$. By \cref{prop:ring-equality}, $R(X,-K_X)\cong R(Y,-K_Y)$, so the anticanonical model
\[
X\cong \Proj R(Y,-K_Y)
\]
is the Stein factorization of the $|-2K_Y|$-map: the morphism $Y\to Q$ factors as $Y\xrightarrow{f} X\xrightarrow{\pi} Q$, where $f$ contracts the eight $(-2)$-curves and $\pi$ is finite of degree $2$.

The branch divisor computation on $\mathbb F_2$ follows from the standard theory of degree-one del Pezzo surfaces \cite[Prop.~2.2]{AscherBejleri21}, \cite{HW81}: after resolving the vertex of $Q$ via $\rho\colon\mathbb F_2\to Q$, the branch locus of $\pi$ pulls back to $B=S+C$ with $C\in |3M|$ reduced and $C\cdot S=0$. The exceptional curve $E_q$ of $\sigma\colon Z\to Y$ is a section of the genus-one fibration on $Z$ and maps to $S$ on $\mathbb F_2$; it is pointwise fixed by the Bertini involution, hence $S$ is a branch component. The class computation
\[
S\cdot C = S\cdot 3M = 3(S^2+2S\cdot F)=3(-2+2)=0,
\qquad
B\sim S+3M=4S+6F=2(2S+3F)
\]
verifies disjointness and the even branch class.
\end{proof}

\begin{lemma}[Local node correspondence for separable double covers]\label{lem:node-correspondence}
Let $U$ be a smooth surface over a field of characteristic different from $2$, and let
\[
\pi\colon V\to U
\]
be a finite separable double cover branched along a reduced divisor $B\subset U$. Let $p\in B$ be a point where $U$ is smooth and $B$ has an ordinary node. Then the unique point of $V$ lying over $p$ is a Du Val singularity of type $A_1$. Conversely, if $B$ is reduced and the point of $V$ over $p$ is of type $A_1$, then $B$ has an ordinary node at $p$.
\end{lemma}

\begin{proof}
Choose local coordinates $(x,y)$ on $U$ centered at $p$. Because $\pi$ is separable and $2$ is invertible, the cover is analytically given by an equation
\[
z^2=f(x,y)
\]
with branch divisor $f=0$. If $B$ has a node at $p$, then after a formal change of coordinates,
\[
f(x,y)=xy+\text{(higher order terms)}
\]
with nondegenerate quadratic part. Completing the square and absorbing higher-order units, the singularity is analytically equivalent to
\[
z^2=xy,
\]
which is the standard $A_1$ singularity. The converse follows from the same normal form: an $A_1$ point of the cover is analytically equivalent to $z^2=uv$, and projecting to the base recovers a reduced branch divisor with ordinary quadratic part $uv=0$, hence an ordinary node.
\end{proof}

\begin{remark}\label{rem:node-on-X}
In our situation the finite double cover is $\pi\colon X\to Q$, not the morphism from the smooth surface $Z$. The eight $A_1$-points of $X$ lie over the smooth locus of $Q$ (and away from the vertex, since the corresponding $(-2)$-curves on $Y$ avoid the anticanonical base point by \cref{lem:basepoint-avoids-minus2}). On $\mathbb F_2$, their images lie away from $S$ and in the smooth locus of the ambient surface, so locally the branch equation is determined by $C$ alone. By \cref{lem:node-correspondence}, the eight $A_1$-points of $X$ correspond to eight ordinary nodes of $C$.
\end{remark}

\subsection{The branch-node bound}

\begin{lemma}[Component partition for divisors in $|3M|$ disjoint from $S$]\label{lem:component-partition}
Let $C\in |3M|$ be a reduced effective divisor on $\mathbb F_2$ with $C\cdot S=0$. Then every irreducible component of $C$ has class $aM$ for some positive integer $a$, and the multiplicities form a partition of $3$.
\end{lemma}

\begin{proof}
Let $D$ be an irreducible component of $C$. Write $D\sim \alpha S+\beta F$ with $\alpha\ge 0$. Since $D$ is effective and irreducible with $D\ne S$, one has $D\cdot S=\beta-2\alpha\ge 0$. But $D$ is a component of $C$ and $C\cdot S=0$, so $D\cdot S\ge 0$. Since every component of $C$ satisfies $D\cdot S\ge 0$ and the total intersection is $C\cdot S=0$, each component satisfies $D\cdot S=0$. Hence $\beta=2\alpha$ and $D\sim \alpha(S+2F)=\alpha M$. The components partition the class $3M$, giving a partition of $3$.
\end{proof}

\begin{lemma}[Node bound for reduced divisors in $|3M|$ on $\mathbb F_2$]\label{lem:node-bound}
Let $C\in |3M|$ be a reduced divisor on $\mathbb F_2$, where $M=S+2F$. Then $C$ has at most six ordinary nodes.
\end{lemma}

\begin{proof}
By \cref{lem:component-partition}, the irreducible components of $C$ partition the class $3M$ into multiples of $M$, so only the partitions $3$, $2+1$, or $1+1+1$ of the coefficient arise. Since
\[
M^2=2,
\qquad
p_a(M)=0,
\qquad
(2M)^2=8,
\qquad
p_a(2M)=1,
\qquad
(3M)^2=18,
\qquad
p_a(3M)=4,
\]
there are only three reduced component partitions to consider.

\medskip
\noindent
\textbf{Partition $3M$.}
If $C$ is irreducible of class $3M$, then its arithmetic genus is $4$, so the number of nodes is at most $4$.

\medskip
\noindent
\textbf{Partition $2M+M$.}
Write $C=C_2+C_1$ with $C_2\in |2M|$ and $C_1\in |M|$. The component $C_2$ has arithmetic genus $1$, hence contributes at most one intrinsic node. The component $C_1$ is smooth rational. Their intersection number is
\[
(2M)\cdot M=4.
\]
Each intersection point of $C_1$ and $C_2$ is at worst an ordinary node of $C_1\cup C_2$ (a tangency or higher contact would produce a singularity worse than a node, reducing the node count). Therefore the total number of nodes is at most
\[
1+4=5.
\]

\medskip
\noindent
\textbf{Partition $M+M+M$.}
Each $M$-component is smooth rational and any two such components meet in
\[
M^2=2
\]
points. Hence the total number of nodes is at most
\[
3\cdot 2=6.
\]
Thus every reduced divisor in $|3M|$ has at most six nodes.
\end{proof}

\begin{theorem}[The pure $8A_1$ basket is impossible]\label{thm:pure8a1-impossible}
There is no rank-one Gorenstein del Pezzo surface in characteristic $3$ with basket $8A_1$.
\end{theorem}

\begin{proof}
Assume such an $X$ exists. By \cref{prop:8a1-degree-one}, it has degree $1$. Let $f\colon Y\to X$ be its crepant resolution. By \cref{prop:deg1-weak-dp}, $Y$ is a degree-one weak del Pezzo surface. By \cref{prop:ring-equality}, the anticanonical ring of $X$ agrees with that of $Y$.

By \cref{prop:branch-model}, the anticanonical model $X\cong\Proj R(Y,-K_Y)$ admits a finite separable double cover $\pi\colon X\to Q\cong \Pbb(1,1,2)$, and after resolving $Q$ to $\mathbb F_2$, the branch divisor has the form $B=S+C$ with $C\in |3M|$ reduced and $C\cdot S=0$.

By \cref{rem:node-on-X}, the eight $A_1$-singularities of $X$ correspond, via \cref{lem:node-correspondence}, to eight ordinary nodes of $C$. Therefore $C$ must have at least eight nodes. But by \cref{lem:node-bound}, a reduced divisor $C\in |3M|$ on $\mathbb F_2$ has at most six ordinary nodes. This contradiction proves that $8A_1$ cannot occur.
\end{proof}

\section{Proof of the main theorem}

\begin{proof}[Proof of \cref{thm:main}]
Assume for contradiction that $X$ is a rank-one klt del Pezzo surface in characteristic $3$ with $N(X)=8$. By \cref{prop:final-baskets}, its basket is one of the five candidate baskets listed there. The three exotic baskets are impossible by \cref{thm:exotic}. The mixed basket is impossible by \cref{thm:mixed-impossible}. The pure Gorenstein basket $8A_1$ is impossible by \cref{thm:pure8a1-impossible}. Thus no rank-one klt del Pezzo surface in characteristic $3$ has eight singular points. Therefore $N(X)\le 7$.
\end{proof}

The public repository accompanying this paper is organized in two layers. The first layer---the scripts and result files cited in \cref{app:repo}---contains the exact computer-assisted inputs used in the final proof. The second layer records older exploratory branches that are not logically required by the theorem but are retained for provenance and for future machine-assisted work. In particular, the repository preserves the earlier ambient/deformation lanes, quasi-elliptic probes, wild additive-action scans, and the matroid-style structural screens that were useful during search even though they do not appear as formal hypotheses in the final argument.

\appendix

\section{The basket-reduction computation}\label{app:basket}
This appendix records the exact data behind \cref{prop:final-baskets}. The point is not to reproduce every line of the scripts in prose, but to expose enough structure and exact counts that a referee can audit the reduction without reverse-engineering the repository.

\subsection{Local cyclic types retained by the \texorpdfstring{$N=8$}{N=8} screen}
The local screen retains the $17$ cyclic quotient types listed in \cref{tab:localtypes}. For each type we record the Hirzebruch--Jung chain, the exceptional rank, the local discrepancy correction $\delta_P:=K_X^2-K_Y^2$ contributed by that singularity, and the determinant of the exceptional lattice.

\begin{longtable}{>{$}c<{$}>{$}c<{$}c c c c}
\caption{Local cyclic quotient types used in the exact basket scan.}\label{tab:localtypes}\\
\toprule
r & q & chain & $\ExcRank$ & $\delta_P$ & det \\
\midrule
\endfirsthead
\toprule
r & q & chain & $\ExcRank$ & $\delta_P$ & det \\
\midrule
\endhead
2 & 1 & [2] & 1 & 0 & 2 \\
3 & 1 & [3] & 1 & 1/3 & 3 \\
4 & 1 & [4] & 1 & 1 & 4 \\
5 & 1 & [5] & 1 & 9/5 & 5 \\
6 & 1 & [6] & 1 & 8/3 & 6 \\
7 & 1 & [7] & 1 & 25/7 & 7 \\
8 & 1 & [8] & 1 & 9/2 & 8 \\
9 & 1 & [9] & 1 & 49/9 & 9 \\
10 & 1 & [10] & 1 & 32/5 & 10 \\
11 & 1 & [11] & 1 & 81/11 & 11 \\
3 & 2 & [2,2] & 2 & 0 & 3 \\
5 & 2 & [3,2] & 2 & 2/5 & 5 \\
8 & 3 & [3,3] & 2 & 1 & 8 \\
7 & 2 & [4,2] & 2 & 8/7 & 7 \\
11 & 3 & [4,3] & 2 & 20/11 & 11 \\
9 & 2 & [5,2] & 2 & 2 & 9 \\
11 & 2 & [6,2] & 2 & 32/11 & 11 \\
\bottomrule
\end{longtable}

\subsection{From local data to a finite basket list}
The scripts implement the following exact finite procedure.
\begin{enumerate}[label=\textup{(\alph*)}]
\item Compute the local invariants in \cref{tab:localtypes}: Hirzebruch--Jung chain, exceptional rank, determinant, local discrepancy correction, and Cartier index.
\item Form all unordered $8$-point multisets from those local types.
\item For each multiset $\mathcal B$, compute the candidate degree
\[
K_X^2(\mathcal B)=9-\sum r_i+\sum \delta_i
\]
as in \cref{lem:basket-screens}, and discard every basket with nonpositive degree.
\item Impose the coarse square-discriminant condition
\[
|A_R|m^2K_X^2\in (\mathbf Z_{>0})^2.
\]
\item On the low-determinant window, run the exact cyclic-complement search in the discriminant group.
\item Replace the provisional quadratic-form filter by the corrected odd-lattice screen on the discriminant bilinear form and anticanonical complement.
\end{enumerate}
At no stage is there any numerical approximation; all quantities are integral or rational and every search space is finite.

\subsection{Exact counts}
The exact enumeration and screening pipeline is:
\begin{enumerate}[label=\textup{(\arabic*)}]
\item Enumerate all $8$-point multisets built from the $17$ local types in \cref{tab:localtypes}. This gives
\[
735{,}471
\]
multisets.
\item Retain those with positive candidate $K_X^2$. This leaves
\[
732{,}878
\]
multisets.
\item Impose the coarse square-discriminant budget. This leaves
\[
4{,}894
\]
multisets.
\item Resolve the small-discriminant window exactly. In the corrected pipeline, this stage resolves $45$ baskets in total: $24$ survive the preliminary exact discriminant stage and $21$ are eliminated there. Earlier provisional runs had left two of these $45$ temporarily unresolved, but the corrected pipeline used in this paper resolves both of them negatively.
\item Replace the provisional quadratic-form screen by the corrected odd-lattice screen on the discriminant bilinear form and anticanonical complement. This leaves exactly the five baskets of \cref{prop:final-baskets}.
\end{enumerate}

Two historical corrections are worth recording because the present appendix deliberately uses only the corrected pipeline. First, an earlier quadratic-form implementation falsely rejected the pure $8A_1$ basket because of a bilinear-pairing normalization error (the old v75 rejection). Second, the v78 provisional odd-lattice screen had to be replaced by the corrected v79 bilinear/anticanonical screen. The present paper uses only the corrected output after those repairs.

\section{Computational ingredients used later in the paper}\label{app:later-computations}
For convenience we record the exact finite counts used in the later sections.

\subsection*{Exotic baskets}
For the normalized exotic blocks of \cref{prop:exotic-normalization}, the corrected orthogonal real-root counts are:
\[
56 \text{ roots for } \tfrac17(1,1),
\qquad
48 \text{ roots for } \tfrac17(1,2),
\qquad
32 \text{ roots for } \tfrac1{11}(1,2).
\]
The repository contains the explicit orthogonality graphs and exact clique computations used in \cref{lem:exotic-root-types}. In the two chain cases the maximal size of a pairwise orthogonal subset is $4$.

\subsection*{Mixed basket}
For the mixed basket:
\begin{itemize}
\item the polarization solve leaves exactly two orbits $L_1$ and $L_2$;
\item for $L_1$, there are $90$ roots with $L_1\cdot r=0$, $210$ roots with $L_1\cdot r=-2$, and $30{,}240$ transformed noncanonical five-cliques, none compatible with an $A_2+A_2+A_1$ subsystem;
\item for $L_2$, there are $128$ roots with $L_2\cdot r=0$, $177$ roots with $L_2\cdot r=-2$, $13{,}440$ canonical $A_2+A_2+A_1$ subsystems split into two Weyl orbits of sizes $3360$ and $10080$, and $5600$ transformed noncanonical five-cliques, all compatible only with the $3360$-orbit.
\end{itemize}
After grouped reduction, there are exactly $16$ grouped five-cliques and exactly $7$ symmetry orbits.

\section{Repository map}\label{app:repo}
The repository is intended to make the computer-assisted parts of the proof auditable and reproducible, for both human referees and machine-assisted verification. It is therefore organized in two layers.

\begin{enumerate}[label=\textup{(\roman*)}]
\item The \emph{final proof layer} contains the scripts, JSON summaries, and notes that are actually cited in the paper.
\item The \emph{exploratory layer} preserves earlier branches that were useful for search, correction, and provenance, but are not load-bearing inputs to the theorem unless explicitly cited.
\end{enumerate}

\subsection*{Final proof layer}
The final theorem depends on the following artifacts.
\begin{enumerate}[label=\textup{(\Alph*)}]
\item The basket reduction of \cref{prop:final-baskets} is recorded in the scripts and outputs corresponding to the v74--v79 chain.
\item The exotic-basket normalization and root-count computations are recorded in the v94 scripts and output summaries.
\item The mixed-basket polarization reduction and the global $L_2$ normalization theorem are recorded in the v95, v96, v97, and v98 artifacts.
\item The pure $8A_1$ branch-node computation is recorded in the v72 and v99 artifacts.
\end{enumerate}

\subsection*{Exploratory layer}
The public repository also preserves three older search lanes because they are useful for understanding the evolution of the argument and for future machine-assisted continuation.
\begin{enumerate}[label=\textup{(E\arabic*)}]
\item \textbf{Ambient/deformation and Picard-merge lane.} This directory contains earlier projective and higher-dimensional search infrastructure, including matroid-style structural screens, compiled finite-field backends, and the associated handoff notes. These files are not used as theorem inputs in the present paper.
\item \textbf{Quasi-elliptic/global-probe lane.} This directory records later ambient and quasi-elliptic scans that helped close or deprioritize several branches before the final blowup-lattice proof stabilized.
\item \textbf{Wild $\alpha_3$/Artin--Schreier lane.} This directory preserves the additive-action and Artin--Schreier no-go ledgers and associated search scripts. They are historically important because they explain why several characteristic-$3$ loopholes were investigated and then abandoned.
\end{enumerate}

A future reader should therefore use the following rule: if a file lives in the exploratory layer and is not explicitly cited in the body of the paper, then it is preserved for provenance only and not as a theorem dependency.

The GitHub-ready archive accompanying this draft preserves both the paper files and the machine-readable computation summaries. The public repository URL for this archive is \repourl, which points to the reproducibility repository corresponding to this draft.

\begin{thebibliography}{99}

\bibitem{AscherBejleri21}
K.~Ascher and D.~Bejleri,
\emph{Moduli of double covers and degree one del Pezzo surfaces},
Algebraic Geometry \textbf{8} (2021), no.~3, 291--330.

\bibitem{CasasAlvero00}
E.~Casas-Alvero,
\emph{Singularities of Plane Curves},
London Math.\ Soc.\ Lecture Note Ser.~\textbf{276}, Cambridge Univ.~Press, 2000.

\bibitem{Demazure80}
M.~Demazure,
\emph{Surfaces de Del Pezzo I--V},
in \emph{S\'eminaire sur les Singularit\'es des Surfaces},
Lecture Notes in Mathematics \textbf{777}, Springer, 1980.

\bibitem{HW81}
F.~Hidaka and K.~Watanabe,
\emph{Normal Gorenstein surfaces with ample anti-canonical divisor},
Tokyo J. Math. \textbf{4} (1981), no.~2, 319--330.


\end{thebibliography}

\end{document}
